---
title: "Reproducible Research Assignment"
author: "Author: Sèche"
date: "First version: 5.3.2017"
output: html_document
---

### Preliminary Steps
Set knitr and R options
```{r setup1}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(digits = 2)
```
Load necessary packages

```{r setup2}
library(dplyr)
library(ggplot2)
```

Unzip and read the data
```{r, }
unzip(zipfile="repdata%2Fdata%2Factivity.zip")  
dat <- read.csv("activity.csv")
```
### Plot histogram before imputing missing values
Compute sum of steps by day and plot histogram
```{r, }
dat1 <- dat %>% group_by(date) %>% summarize(sumstepsbyday = sum(steps))
ggplot(dat1, aes(sumstepsbyday)) + geom_histogram() + xlab("Steps per Day") + ylab("Nr. of Days")
```

### Mean and Median
Compute the avg. nr. of steps and the median sum of steps per day
```{r, }
mean_raw <- mean(dat1$sumstepsbyday, na.rm = TRUE)
median_raw <- median(dat1$sumstepsbyday, na.rm = TRUE)
mean_raw
median_raw
```
### Plot time series of activity
Compute the avg. nr. of steps per day and plot as time series
```{r, }
dat2 <- dat %>% group_by(interval) %>% summarize(avgstepsbyday = mean(steps, na.rm = TRUE))
ggplot(dat2, aes(x = interval, y = avgstepsbyday, group = 1)) +
        geom_line() + geom_smooth() + xlab("Time") + ylab("Avg. Steps per Day")
```

### Retrieving the interval with most activity 
Get the row with the max avg. nr. of steps
```{r, }
maxactiveinterval <- dat2[which.max(dat2$avgstepsbyday), ]
maxactiveinterval
```
The interval with most activity is `r maxactiveinterval[1, 1]` with `r  maxactiveinterval[1, 2]` average steps per day.

### Count nr. of rows with NA
```{r, }
na_count <- sum(is.na(dat$steps))
na_count
```
There are `r na_count` rows with missing values.

### Imputing missing values 
Replace NA with average steps of corresponding interval of other days
```{r, }
dat3 <- dat %>%
        group_by(interval) %>% 
        mutate_each(funs(replace(., which(is.na(.)),
                                 mean(., na.rm = TRUE))))
```

### Plot histogram after imputing missing values
Compute the sum of steps per day and plot histogram after missing data treatment
```{r, }
dat4 <- dat3 %>% group_by(date) %>% summarize(sumstepsbyday = sum(steps))
ggplot(dat4, aes(sumstepsbyday)) + geom_histogram() + xlab("Steps per Day") + ylab("Nr. of Days")
```

### Compute the avg. nr. of steps and the median nr. of steps per day after missing data treatment
```{r, }
dat5 <- dat3 %>% group_by(date) %>% summarize(sumstepsbyday = sum(steps))
mean_clean <- mean(dat5$sumstepsbyday, na.rm = TRUE)
median_clean <- median(dat5$sumstepsbyday, na.rm = TRUE)
mean_clean
median_clean
```
The mean didnt' change since we imputed the missing data with mean values of all other days. The median has, however, changed slighlty: From `r mean_raw`(before missing data treatment) to `r mean_clean`(after treatment).

### Plot time series panel by weekday
Convert "date" to date format
```{r, }
dat3$date <- as.Date(dat3$date)
```

Compute dummy variable for weekday and weekend
```{r, }
dat3$weekday <- ifelse(weekdays(dat3$date) == "Samstag" | weekdays(dat3$date) == "Sonntag",
                       "weekend", "weekday")
```

Compute avg. nr. of steps per interval and weekday and plot as time series per weekday (panel) 
```{r, }
dat6 <- dat3 %>% group_by(weekday, interval) %>% summarize(avgbyweekday = mean(steps))
ggplot(dat6, aes(interval, avgbyweekday)) + geom_line() + facet_grid(weekday ~.) + 
        xlab("Interval") + ylab("Avg. Nr. of Steps")
```


